Parsing Postfix log files

### 2008-04-29 doc/logparser-other-parsers-review.tex
Tested Splunk.

Splunk aims to index all an organisation's logs, providing a centralised
view capable of searching and correlating diverse log sources.  The web
interface supports complicated searches, providing statistics and graphs in
real time, a facility not provided by \parsername{} (report generation has
been deferred to a subsequent program).  Saved searches\footnote{The author
was unable to save searches, though that may have been due to limitations
in the free version.} can be run periodically and the results emailed to a
recipient or sent to a shell script, which presumably can publish the
results as required (though possibly without the graphs and detailed
statistics); \parsername{} provides the database and leaves it to the user
to utilise it in any way, whenever they want, with no restrictions on usage
of the data.  The interface is optimised for interactive rather than
automated queries and it does not appear to be possible to write
independent tools to utilise the Splunk database.  Many types of reports
are available, though most are variations of a bar or pie chart, with the
exception of bubble and heatmap graphs.  It is very easy to drill down
through the graphs to extract a portion of the data (e.g.\ select the hour
with the largest number of events, then select a particular host, and
finally a specific address), though it is not possible to search on partial
words.  The searches are quite fast, though as the free version has a limit
on the amount of data indexed per day the volume of data to be searched was
merely 45MB\@; the cheapest licensed version costs \$6000, and still limits
the volume of data indexed per day.

In the specific case of parsing Postfix logs, Splunk extracts some standard
fields: to and from addresses, HELO hostname, date, host the logs were
collected from, and protocol.  It parses the standard syslog fields at the
beginning of the line, and extracts any \texttt{key=<value>} pairs
occurring after the standard syslog prologue; these pairs are the fields
listed above.  Searches can be based on the extracted fields, and all text
in the line is also available for searching.  \parsername{} extracts
noticeably more data, though it does not make the full text of the line
available, and the full power of \SQL{} is available when searching,
allowing the user to search on arbitrarily complicated conditions.

Splunk is a generic tool, so it lacks any Postfix specific support over and
above extracting the \texttt{key=<value>} fields; most importantly it makes
no attempt to correlate log lines by queueid or \pid{}, nor to handle any
of the myriad complications discussed in this document.

\url{http://www.splunk.com/} \newline (Last checked 2007/09/25.)

>>>>

Splunk aims to index all an organisation's logs, providing a centralised
view capable of searching and correlating diverse log sources.  The web
interface supports complicated searches, providing statistics and graphs in
real time, a facility not provided by \parsername{} (report generation has
been deferred to a subsequent program).  Saved searches\footnote{The author
was unable to save searches, though that may have been due to limitations
in the free version.} can be run periodically and the results emailed to a
recipient or sent to a shell script, which presumably can publish the
results as required (though possibly without the graphs and detailed
statistics); \parsername{} provides the database and leaves it to the user
to utilise it in any way, whenever they want, with no restrictions on usage
of the data.  The interface is optimised for interactive rather than
automated queries and it does not appear to be possible to write
independent tools to utilise the Splunk database; there are additional
reports available on \url{http://www.splunkbase.com/}, but the author was
unable to find any --- every category was empty, even those which the
interface claimed had reports available.  Many types of reports are bundled
with the software, though most are variations of a bar or pie chart, with
the exception of bubble and heatmap graphs.  It is very easy to drill down
through the graphs to extract a portion of the data (e.g.\ select the hour
with the largest number of events, then select a particular host, and
finally a specific address), though it is not possible to search on partial
words.

The web interface is quite attractive and easy to use when searching, but
as an administrator it seems unnecessarily difficult to perform simple
tasks.  When testing it took roughly 30 minutes to add a single log file to
be indexed (log files must be indexed before searching), with the downside
that the log file was copied into a spool directory, doubling the disk
space usage.  The most suitable test would be to index all log files in a
particular directory, but after three hours, numerous futile attempts, and
reading all available documentation the author admitted defeat.  Using the
\CLI{} was more successful: invoking the command \newline \tab{}
\texttt{splunk find logs }\textit{log-directory\/}\newline added 40 of the
\numberOFlogFILES{} log files to the queue for indexing.  Repeated attempts
enqueued the same 40 log files, without explanation as to why the others
were excluded.\footnote{The log files appear to have only been indexed
once; presumably Splunk keeps track of the files it has indexed and
discards requests to index files for a second time.  This may or may not be
a useful feature for \parsername{}.} There did not appear to be an option
to order the log files to ensure they would be processed in the order they
were created.  Subsequently the author was successful in adding a single
file at a time using the \CLI{}; a simple loop to add all desired log files
was sufficient to index all files.  Splunk will periodically check all
those files for updates unless they are manually removed from its list;
this may or may not be useful behaviour.  Splunk did not appear to have any
difficulty in parsing the log files, once instructed to do so.  All
searches performed using the indexed data returned reasonable results.

Splunk supports reading compressed log files without any configuration by
the user.  The free version of Splunk limits the volume of data indexed per
day to 500MB, though a trial Enterprise licence is available which allows
indexing of up to 5GB of data per day.  The cheapest licensed version costs
\$5000 (plus \$1000 support), and still limits the volume of data indexed
per day to 500MB.

In the specific case of parsing Postfix logs, Splunk extracts some standard
fields: to and from addresses, HELO hostname, date, host the logs were
collected from, and protocol.  It parses the standard syslog fields at the
beginning of the line, and extracts any \texttt{key=<value>} pairs
occurring after the standard syslog prologue; these pairs are the fields
listed above.  Searches can be based on the extracted fields, and all the
text in the line is also available for searching.  \parsername{} extracts
noticeably more data, though it does not make the full text of the line
available, and the full power of \SQL{} is available when searching,
allowing the user to search on arbitrarily complicated conditions.

Splunk is a generic tool, so it lacks any Postfix specific support over and
above extracting the \texttt{key=<value>} fields; most importantly it makes
no attempt to correlate log lines by queueid or \pid{}, nor to handle any
of the myriad complications discussed in this document.  Some additional
Postfix reports are supposedly available at
\url{http://www.splunkbase.com/}, but the author was unable to find them
(or any other reports).

\url{http://www.splunk.com/} \newline (Last checked 2008/04/29.)
